An Application of Graph Theory to MDAO Problem Formulation

### 2012-10-12 intro.tex
added in a new reference, fixed two others

    As the size and complexity of engineering systems grow, the time and expense for setting up 
    analysis models grow with them. Multidisciplinary Design Analysis and Optimization (MDAO)
    frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter have enabled a new level of analysis tool integration 
    and paved the way for models with more analysis tools and increasing numbers of multidisciplinary couplings. 
    Such complex models present a distinct challenge for proper implementation of any given solution strategy.

    We propose the use of graph-based problem formulation that generally describes a
    given problem, supports algorithmic manipulation to find beneficial solutions strategies, and enables automated 
    implementation of those strategies inside an MDAO framework. In order to serve that purpose, the graph includes the following information: 
    \begin{itemize}
       \item Discipline Analyses 
       \item Discipline State Variables and Residuals
       \item Design Variables
       \item Constraints
       \item Objective or Objectives
       \item Coupling Constraints
       \item Parameters
    \end{itemize}
    Notably absent from the preceding list are any kind of solvers, optimizers, or other iterative solution finding tools. 
    At its most basic, a problem formulation includes only information about what is being sought after in a given problem or what the
    goals of a given problem are. It need not contain any information about solution paths or strategies to reach those goals.

@@

    The FPF in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} and Eq. \ref{eqn:simple2} because it has 
    two constraints which both must be met. The presence of both of these constraints fully decouples the problem so that 
    either $A$ or $B$ could be run first or both could be run simultaneously. By removing either constraint and replacing 
    it with a direct dependence between the two analyses you could regain the earlier two problem formulations.

@@

    An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Michelena and Papalambros. 
    FDT represents the relationship between functions, including objectives and constraints, and specific variables that affect 
    them\cite{Michelena1997}. Similar to DSM, FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, 
    however, the graph is undirected and nodes can represent analysis tools, objectives, 
    or constraints. Edges between nodes represent a dependence on the same 
    variable. Wagner and Papalambros made use of the FDT to solve a graph partitioning problem that yielded 
    more efficient optimization problem decompositions\cite{Wagner1993}. While FDT succeeds at capturing the 
    information about objectives and constraints, it neglects the data dependency captured by DSM. For instance, 
    although we know from the FPF in Eq. \ref{eqn:simple_fpf} that the objective, $f$, is dependent on the 
    output of analysis $A$, you could not determine that from the FDT in Fig. \ref{fig:FDT_simple} alone.

    [TODO: There is more recent research to cite here. e.g. Allison's Work.]

@@

    \begin{figure}
        \begin{center}
        \includegraphics[height=.25\textheight]{XDSM/simple}
        \caption{XDSM for Eqn. \ref{eqn:simple}, with a gauss-siedel iteration and MDF solution architecture. \label{fig:XDSM_simple}}
        \end{center}
    \end{figure}

    Although XDSM captures part of the functional aspects of FDT, it leaves out the variables themselves. 
    As a result, all variable information is aggregated 
    so that for the problem from Eq. \ref{eqn:simple_fpf} you can say that $A$ depends on $B$ or vice versa, but you can't identify which 
    individual variables are interacting in the dependency cycle. Without the detailed variable information, you can't construct the compatibility 
    constraints necessary to implement the problem. Additionally, XDSM requires the use of solver and optimizer blocks to represent 
    the relationship between design variables and objectives/constraints. By introducing solver or optimizer blocks XDSM automatically provides
    some kind of solution strategy. The XDSM for Eqn. \ref{eqn:simple} is given in Figure \ref{fig:XDSM_simple}. 
    This diagram is shown with an assumed gauss-siedel iteration scheme and a MDF solution architecture. 
    Hence XDSM is too specific for use with a fundamental problem formulation.

>>>>

        As the size and complexity of engineering systems grow, the time and expense for setting up 
        analysis models grow with them. Multidisciplinary Design Analysis and Optimization (MDAO)
        frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter have enabled a new level of analysis tool integration 
        and paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. 
        Such complex models present a distinct challenge for proper implementation of any given solution strategy.

        We propose the use of graph-based problem formulation that generally describes a
        given problem, supports algorithmic manipulation to find beneficial solutions strategies, and enables automated 
        implementation of those strategies inside an MDAO framework. In order to serve that purpose, the graph must include the following information: 
        \begin{itemize}
           \item Discipline Analyses 
           \item Discipline State Variables and Residuals
           \item Design Variables
           \item Constraints
           \item Objective or Objectives
           \item Coupling Constraints
           \item Parameters
        \end{itemize}
        Notably absent from the preceding list are any kind of solvers, optimizers, or other iterative solution finding tools. 
        At its most basic, a problem formulation includes only information about what is being sought after in a given problem or what the
        goals of a given problem are. It need not contain any information about solution paths or strategies to reach those goals.

@@

        The FPF in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} and Eq. \ref{eqn:simple2} because it has 
        two constraints which both must be met. The presence of both of these constraints fully decouples the problem so that 
        either $A$ or $B$ could be run first or both could be run simultaneously. By removing either constraint and replacing 
        it with a direct dependence between the two analyses you could regain the earlier two problem formulations. Alexandrov
        and Lewis demonstrated the value of a more modular approach to problem formulation because it enables one to transition between different 
        MDAO solution strategies depending on the specifics of the problem\cite{Alex2000}.

@@

        An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros. 
        FDT represents the relationship between functions, including objectives and constraints, and specific variables that affect 
        them\cite{Wagner1993}. Similar to DSM, FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, 
        however, the graph is undirected and nodes can represent analysis tools, objectives, 
        or constraints. Edges between nodes represent a dependence on the same 
        variable. Michelena and Papalambros made use of the FDT to solve a graph partitioning problem that yielded 
        more efficient optimization problem decompositions\cite{Michelena1997}. While FDT succeeds at capturing the 
        information about objectives and constraints, it neglects the data dependency captured by DSM. For instance, 
        although we know from the FPF in Eq. \ref{eqn:simple_fpf} that the objective, $f$, is dependent on the 
        output of analysis $A$, you could not determine that from the FDT in Fig. \ref{fig:FDT_simple} alone. This make FDT 
        applicable for partitioning decisions, but not for ordering problems.

@@

        \begin{figure}
            \begin{center}
            \includegraphics[height=.25\textheight]{XDSM/simple}
            \caption{XDSM for Eq. \ref{eqn:simple}, with a gauss-siedel iteration and MDF solution architecture. \label{fig:XDSM_simple}}
            \end{center}
        \end{figure}

        Although XDSM captures part of the functional aspects of FDT, it leaves out the variables themselves. 
        As a result, all variable information is aggregated 
        so that for the problem from Eq. \ref{eqn:simple_fpf} you can say that $A$ depends on $B$ or vice versa, but you can't identify which 
        individual variables are interacting in the dependency cycle. Without the detailed variable information, you can't construct the compatibility 
        constraints necessary to implement the problem. Additionally, XDSM requires the use of solver and optimizer blocks to represent 
        the relationship between design variables and objectives/constraints. By introducing solver or optimizer blocks XDSM automatically provides
        some kind of solution strategy. The XDSM for Eq. \ref{eqn:simple} is given in Figure \ref{fig:XDSM_simple}. 
        This diagram is shown with an assumed gauss-siedel iteration scheme and a MDF solution architecture. 
        Hence XDSM is too specific for use with a fundamental problem formulation.
